\section*{Abstract} 
The results from neutrino oscillation experiments indicate that at
least two neutrinos have mass. However, the value of the masses and
whether neutrinos and anti-neutrinos are identical, \textit{i.e.},
Majorana particles, remain unknown. Neutrinoless double beta decay
experiments can help to improve our understanding in both cases and
are the only method currently possible to tackle the second question.
 
The GERmanium Detector Array (GERDA) experiment, which will search for
the neutrinoless double beta decay of $^{76}$Ge, is currently under
construction in Hall A of the INFN Gran Sasso National Laboratory
(LNGS), Italy. In order to achieve an extremely low background level,
segmented germanium detectors are considered to be operated directly
in liquid argon which serves simultaneously as cooling and shielding
medium.
 
Several test cryostats were built at the Max-Planck-Institut f\"ur
Physik in M\"unchen to operate segmented germanium detectors both in
vacuum and submerged in cryogenic liquid. The performance and the
background discrimination power of segmented germanium detectors were
studied in detail. It was proven for the first time that segmented
germanium detectors can be operated stably over long periods submerged
in a cryogenic liquid. It was confirmed that the segmentation scheme
employed does well in the identification of photon induced background
and demonstrated for the first time that also neutron interactions can
be identified.
 
The C++ Monte Carlo framework, MaGe (Majorana-GERDA), is a joint
development of the Majorana and GERDA collaborations. It is based on
GEANT4, but tailored especially to simulate the response of ultra-low
background detectors to ionizing radiation. The predictions of the
simulation were verified to be accurate for a wide range of
conditions. Some shortcomings were found and corrected.
 
Pulse shape analysis is complementary to segmentation in identifying
background events.  Its efficiency can only be correctly determined
using reliable pulse shape simulations.  A fully functional pulse
shape simulation package was developed to augment the MaGe
package. The simulation was verified using data taken with the first
segmented prototype detector for GERDA. This work also led to a
considerable improvement in the understanding of segmented germanium
detectors.
 
\clearpage 
 
\section*{Zusammenfassung} 
Die Ergebnisse von Neutrinooszillationsexperimenten zeigen, dass
mindestens zwei Neutrinos eine endliche Masse haben. Die absolute
Massenskala ist jedoch nicht bekannt. Ungel\"ost ist auch die Frage,
ob das Neutrino sein eigenes Antiteilchen ist, \textit{i.e.}, ob
Neutrinos Majoranateilchen sind. Eine m\"ogliche Beobachtung von
neutrinolosem Doppelbetazerfall k\"onnte zur Feststellung der
Massenskala beitragen und ist im Moment die einzige realisierbare
M\"oglichkeit, die Frage nach der Natur der Neutrinos zu kl\"aren.
  
Das GERmanium Detector Array (GERDA) f\"ur die Suche nach
neutrinolosem Doppelbetazerfall von $^{76}$Ge wird derzeit in der
Halle A des ``INFN Gran Sasso National Laboratory (LNGS)'' in Italien
aufgebaut. Um ein extrem niedriges Untergrundniveau zu ereichen,
werden segmentierte Germaniumdetektoren direkt in fl\"ussigem Argon,
das gleichzeitig als K\"uhl-und Abschirmmedium dient, betrieben.

Mehrere Testkryostaten wurden f\"ur den Betrieb segmentierter
Germaniumdetektoren in Vakuum und kryogener Fl\"ussigkeit am
Max-Planck-Institut f\"ur Physik in M\"unchen entwickelt. Es wurde zum
ersten Mal gezeigt, dass segmentierte Germaniumdetektoren \"uber lange
Zeit stabil in einer kryogenen Fl\"ussigkeit betrieben werden
k\"onnen. Die M\"oglichkeiten, Untergrundereignisse in segmentierten
Germaniumdetektoren zu identifizieren, wurden im Detail untersucht.
Dabei wurde betst\"atigt, dass geeignete Segmentierung die
Identifikation von photoninduzierten Ereignissen erm\"oglicht und zum
ersten Mal gezeigt, dass auch neutroninduzierte Ereignisse
identifiziert werden k\"onnen.

Das C++ Monte Carlo Paket MaGe (Majorana-GERDA) ist ein
Gemeinschaftprojekt der Majorana und GERDA Kollaborationen. Es basiert
auf GEANT4, ist aber zurechtgeschnitten auf die Simulation der
Wechselwirkungen ionisierender Strahlung mit Detektoren f\"ur
Experimente in einer Umgebung mit extrem niedriger Radioaktivit\"at.
Die Simulationen wurden mit Daten \"uberpr\"uft. Daten und Simulation
stimmen f\"ur viele Anwendungsf\"alle gut \"uberein.

Pulsformanalyse kann Segmentierung bei der Identifikation von dem
Untergrund zuzurechnenden Ereignistopologien erg\"anzen. Zur
Ermittlung der Effizienz von Pulsformanalysen wird eine exakte
Pulsformsimulation ben\"otigt. Ein komplettes Pulsformsimulationspaket
wurde im Rahmen diesr Arbeit entwickelt. Die simulierten Pulse wurden
mit Messungen am ersten segmentierten Prototypdetektor verglichen und
den Daten erfolgreich angepasst. Dabei wurde auch das Verst\"andnis
der Detektoren signifikant verbessert.
 
\clearpage 
 
%%% Local Variables: 
%%% mode:latex 
%%% TeX-master: "thesis" 
%%% End: 
