\chapter{Acknowledgments}
\label{cha:ack}

First of all, I would like to thank Prof. Dr. Allen Caldwell for allowing me to work in his group. Many thanks to Dr. Iris Abt for having chosen me in the second IMPRS Ph.D. student selection and her encouragement all along my three years of studies despite of my horrible English and stupidity. Thanks a lot for the help from Dr. Iris Abt and Dr. B\'ela Majorovits with my studies in their groups. I can hardly forget the paper reading with them in the beer garden.  And thank all my bosses for their valuable guidance in physics and kind considerations in my personal life.

I appreciate the big help from Xiang. He explained to me patiently all the details about our work during my first days here when I hardly spoke any English. I enjoyed the cooperation with Kevin and Xiang in the background study, with Daniel in the pulse shape simulation and with Dan, Andreas and Gregor in LabVIEW programming. Thanks a lot for their suggestions to my parts of work and selfless sharing with me their own efforts. I received big helps from Jens with statistics and C++ problems. I learned many from Jozsef about basic electronics. Luciano, Jason, Mike and all the members of the MC groups in both GERDA and Majorana collaborations helped a lot with my work on background and pulse shape simulations. I would like to thank them all. My thanks also go to all our interns and summer students for their helps on our daily lab work.

Thank all the members of the Chinese book-reading club and all the friends and colleagues at our institute for the happy times we spent together.

And finally, big thanks to my wife, for her love, patience and support.


%%% Local Variables:
%%% mode:latex
%%% TeX-master: "thesis"
%%% End:
