Neutrinoless double beta decay is an extremely rare process even if it does exist. Requirements, which are crucial to increase the sensitivity of $0\nu\beta\beta$ decay experiments, are summarized. The ``Pros and Cons'' of different experimental approaches are discussed based on these requirements.

\section{Sensitivity}
\label{sec:sensi}
The number of observed $0\nu\beta\beta$ decay events $N_{s}$ within the measuring time $t$, can be calculated as
\begin{equation}
  \label{eq:gerda:ns}
  N_{s} = M \cdot \kappa \cdot \frac{N_{A}}{M_{A}} \cdot \epsilon \cdot (1 - e^{-t/\tau}) \approx M \cdot \kappa \cdot \frac{N_{A}}{M_{A}} \cdot \epsilon \cdot \frac{t}{\tau},
\end{equation}
where, $M$ is the total mass of the source material, $\kappa$ is the mass fraction of the isotope under study, $N_{A}$ is Advogadro's number, $M_{A}$ is the atomic mass of the isotope, $\epsilon$ is the signal detecting efficiency, and $\tau$ is the mean lifetime of the decay. Since the measuring time $t$ is much shorter than the mean lifetime $\tau$, $(1 - e^{-t/\tau})$ is approximated as $t/\tau$. The half lifetime, $T^{0\nu}_{1/2}$, is then
\begin{equation}
  \label{eq:gerda:thalf}
  T^{0\nu}_{1/2} = \ln2 \cdot \tau \approx \ln2 \cdot M \cdot \kappa \cdot \frac{N_{A}}{M_{A}} \cdot \epsilon \cdot \frac{t}{N_{s}}.
\end{equation}
The number of background events within the measuring time $t$ and within the energy window of interest $\Delta E$ is 
\begin{equation}
  \label{eq:gerda:nb}
  N_{b} = b \cdot M \cdot t \cdot \Delta E,
\end{equation}
where $b$ is the background index given in per kilogram of source material per measuring year and per keV. If $N_{s}$ is smaller than the standard fluctuation expected for $N_{b}$, \textit{i.e.} $N_{s}<\sqrt{N_{b}}$, the signal cannot be extracted. In this case the relation
\begin{equation}
  \label{eq:gerda:thalfb}
  T^{0\nu}_{1/2} > \ln2 \cdot M \cdot \kappa \cdot \frac{N_{A}}{M_{A}} \cdot \epsilon \cdot \frac{t}{\sqrt{N_{b}}} = \ln2 \cdot \kappa \cdot \frac{N_{A}}{M_{A}} \cdot \epsilon \sqrt{\frac{M t}{b \Delta E}}
\end{equation}
can be used to set a lower limit on the half lifetime. Combined with Eq.~\ref{eq:0nurate}, the following relation can be deduced to set an upper limit on the effective Majorana neutrino mass:
\begin{equation}
  \label{eq:gerda:mbb}
  m_{\beta\beta} < \sqrt{\frac{M_{A}}{\ln2 \cdot \kappa \cdot N_{A} \cdot \epsilon}} \sqrt{\frac{1}{G_{0\nu}(Q,Z)}} \frac{1}{|\mathcal{M}_{0\nu}|} (\frac{b \Delta E}{M t})^{1/4}
\end{equation}
The sensitivity of a
$0\nu\beta\beta$ decay experiment in terms of lifetime or neutrino mass can be estimated based on Eq.~\ref{eq:gerda:thalfb} or \ref{eq:gerda:mbb}, respectively.

\section{Experimental approaches}
\label{sec:exp:appr}
\subsection{General considerations}
\label{sec:gencon}
An analysis of Eq.~\ref{eq:gerda:mbb} provides guidance on how to design a $0\nu\beta\beta$ decay experiment with a good sensitivity. As many of the following requirements should be met:
\begin{itemize}
\item the mass of the source material $M$ should be large;
\item the abundance of the isotope under study should be high, either naturally or by enrichment;
\item the calculation of the nuclear matrix element $|\mathcal{M}_{0\nu}|$ for this isotope should be accurate;
\item the $Q$-value should be large, because $G_{0\nu}(Q,Z) \propto Q^{5}$, the higher the $Q$-value, the fewer lines of the natural radioactive background matter;
\item the signal detecting efficiency should be large;
\item the energy resolution should be good in order to allow a small $\Delta E$;
\item last but not the least, the background level $b$ should be as low as possible.
\end{itemize}
And then patience is needed; data is generally collected over many years.

Except for background suppression techniques, the experimental approaches are mainly determined by the choice of the source material. Table~\ref{tab:gerda:iso} presents a selection of isotopes used or planned to be used to search for $0\nu\beta\beta$ decay. Also listed are their $Q$-values, nuclear matrix elements~\cite{Mut90, Rod07, Sim08, Cau08}, natural abundance $\kappa$ and properties important for the experimental design. Different experimental approaches can be classified into two categories: 1. the source material can be used to produce the detector; 2. the source is not the detector, the decay products need to be detected using equipment around the source.
\begin{table}[htbp]
  \centering
  \caption{A selection of possible source candidates for         $0\nu\beta\beta$ decay experiments. Also listed are their $Q$-values,     nuclear matrix elements, natural abundance $\kappa$ and properties     important for the design of experiments.}
  \label{tab:gerda:iso}
  \begin{minipage}{\linewidth}
    \begin{tabular}{ccccc} \hline Isotope & $Q$ [MeV] &       $\mathcal{M}_{0\nu}$ & $\kappa$ [\%] & Properties \\\hline       $^{48}$Ca & 4.271 & 0.67\footnote{The values are from an ISM         (Interacting Shell Model) calculation in Ref~\cite{Cau08}.         The other $\mathcal{M}_{0\nu}$ values with errors are from         QRPA (Quasi-particle Random Phase Approximation)         calculations~\cite{Rod07}. The errors are from the measurement         of $2\nu\beta\beta$ experiments.} & 0.19 & CaF$_{2}$ \&       CaWO$_{4}$ is a scintillator \\
      $^{76}$Ge & 2.039 & $4.51 \pm 0.17$ & 7.8 & semiconductor \\
      $^{82}$Se & 2.995 & $4.02 \pm 0.15$ & 9.2 & - \\
      $^{96}$Zr & 3.350 & $1.12 \pm 0.03$ & 2.8 & - \\
      $^{100}$Mo & 3.034 & $3.34 \pm 0.19$ & 9.6 & - \\
      $^{116}$Cd & 2.809 & $2.74 \pm 0.19$ & 7.5 &       CdZnTe\footnote{There are other isotopes in the CdZnTe crystal         that could undergo $0\nu\beta\beta$ decay. The rest of them         are $^{70}$Zn with $Q = 1.001$~MeV, $\kappa = 0.62\%$,         $^{114}$Cd with $Q = 0.534$~MeV, $\kappa = 28.7\%$, $^{128}$Te         with $Q = 0.868$~MeV, $\kappa = 31.7\%$ and $^{130}$Te.} is a       semiconductor;\\
      & & & & CdWO$_{4}$ is a scintillator\\
      $^{124}$Sn & 2.287 & $2.11^{a}$ & 5.8 & semiconductor \\
      $^{130}$Te & 2.530 & $3.26 \pm 0.12$ & 35 & TeO$_{2}$ can be       used as bolometer\\
      $^{136}$Xe & 2.480 & $2.11 \pm 0.11$ & 8.9 & active material for       time projection chambers\\
      $^{150}$Nd & 3.367 & $4.74 \pm 0.20$ & 5.6 & could be dissolved       in liquid scintillator\\
    \end{tabular}
  \end{minipage}
\end{table}

\subsection{Source and detector are identical}
\label{sec:exp:sed}
As shown in Table~\ref{tab:gerda:iso}, quite a few $0\nu\beta\beta$ decay candidates have special properties which allow them to be used as detectors. There are advantages to this concept. As the decay electrons do not have to leave the source and reach the detector, the detection efficiency is not limited and the energy resolution of the detector not deteriorated. As a consequence large compact masses are useable limiting the loss of events close to a surface with electrons escaping. The drawback is that such detectors usually have limited capability to reconstruct event topologies and normally only one isotope can be studies.

$^{48}$Ca has the highest $Q$-value among all the candidates. Hence low background from natural radioactivities is expected. It also means a large phase space factor which enlarges the $0\nu\beta\beta$ decay rate for a given Majorana mass. However, until now only few experiments have been carried out because of its low natural abundance. The most stringent limit on the $0\nu\beta\beta$ decay of $^{48}$Ca came from ELEGANT VI~\cite{Oga04} using CaF$_{2}$ scintillator. Two future experiments using CaF$_{2}$ and CaWO$_{4}$ as scintillator, respectively, are CANDLES~\cite{Hir08} and CARVEL~\cite{Zde05}. They aim at a sensitivity in $m_{\beta\beta} <$~(0.04-0.09)~eV.

The search for the $0\nu\beta\beta$ decay of $^{76}$Ge is affected by natural radioactivity due to its low $Q$-value. Enrichment in $^{76}$Ge is also needed in order to overcome the low natural abundance. However, semiconductor detectors made from high purity germanium crystals have been used as gamma spectrometers for years and have an excellent energy resolution. Previous $^{76}$Ge $0\nu\beta\beta$ decay experiments include IGEX~\cite{Aal02} and HdM~\cite{Hei04}. GERDA~\cite{Sch05} Phase I is currently under construction and will be described in detail in the next chapter. The planned future experiments include GERDA Phase II and Majorana~\cite{Gai03, Aal04}. The GERDA and Majorana Collaborations have reached an agreement to share resources and knowledge where appropriate in their parallel development of the two different detector designs. The ultimate goal is to combine the strength of the two collaborations in a future experiment that will employ the best technology for reaching a Majorana neutrino mass sensitivity of below 0.05~eV.

The Cobra experiment~\cite{Zub01, Ell02, Kie03} using a large array of CdZnTe semiconductor detectors is a special case in the ``source = detector'' concept. The CdZnTe crystal contains 5 isotopes which could undergo $0\nu\beta\beta$ decay. Pixellated CdZnTe detectors can be operated as solid-state time projection chambers (TPC) and hence offer tracking capability which allows reconstruction of the event topology. Another advantage of CdZnTe detectors is that they can be operated at room temperature. No large and complicated cooling facility is needed. However, the energy resolution of CdZnTe detectors currently available is not as good as those of germanium detector and TeO$_{2}$ bolometers.

The tungstate CdWO$_{4}$ is similar to CaWO$_{4}$, and can be used as scintillator. A comprehensive comparison between them can be found on page 27 of Ref.~\cite{Avi05}. The CdWO$_{4}$ crystal is less contaminated and has better background/signal discrimination power than the CaWO$_{4}$ crystal. Previous $0\nu\beta\beta$ decay experiments using CdWO$_{4}$ scintillators include the one performed by the Kiev-Florence collaboration in the Solotvina Underground Laboratory since 1989~\cite{Dan00, Dan03} and CAMEO~\cite{Bel00, Bel01}. The CAMEO project also proposes to exploit 1 ton of $^{116}$CdWO$_{4}$ detectors placed in one of the large underground neutrino detectors such as BOREXINO~\cite{Arp08}, SNO or KamLAND. The sensitivity is estimated to be $m_{\beta\beta} < 0.02$~eV.

CUORICINO~\cite{Pre04}, the pilot experiment for CUORE~\cite{Arn04, Ard05}, just released a new upper limit of $m_{\beta\beta} <$~0.19-0.68~eV~\cite{Arn08} using TeO$_{2}$ bolometers. TeO$_{2}$ bolometers have almost the same energy resolution as the germanium detector. $^{130}$Te has the highest natural abundance among all the $0\nu\beta\beta$ decay candidates. CUORE is also trying to use CdWO$_{4}$ as scintillating bolometer~\cite{Gir08}. The scintillating bolometer could provide more parameters helping background rejection, especially from surface contamination. One of the challenges of CUORE is to stabilize the contact between the bolometer and the thermometer.

Xenon can be used as active material in a TPC. The reconstruction of the tracks of the electrons from the beta decay is possible making it easy to discriminate against background induced by natural radioactivity. The main background comes from $2\nu\beta\beta$ due to the limited energy resolution. An experiment was carried out in the Gotthard underground laboratory using TPC filled with xenon gas enriched to 62.5~\% in $^{136}$Xe at a pressure of 5 bar~\cite{Lue98}. Currently, there are two experiments named EXO~\cite{Dan00} and XMASS~\cite{Kim05} proposing to use TPC filled with liquid xenon which can be used as scintillator at the same time.

$^{150}$Nd has the second highest $Q$-value, the largest nuclear matrix element $\mathcal{M}_{0\nu}$ (though the uncertainty is large) and a relatively high natural abundance (enrichment is possible) among all the $0\nu\beta\beta$ decay candidates. In a follow-up experiment to SNO, called SNO+~\cite{Zub07}, the old SNO infrastructure will be filled with Nd-loaded liquid scintillator instead of D$_{2}$O. Although the energy resolution of the detector will not be as good as that of other existing experiments, the mass that could be suspended in the scintillator is very large (0.1\% load of natural Nd corresponds to 56~kg of $^{150}$Nd). Based on preliminary simulations, loading with enriched Nd at 0.1\% level SNO+ would have the sensitivity of $m_{\beta\beta} \sim 0.03$~eV.

\subsection{Source and  detector are not identical}
\label{sec:exp:sued}
The advantages of using external detectors are (1) the source material could be changed so that several $0\nu\beta\beta$ decay candidates could be investigated in the same experiment; (2) using tracking devices the event topology could be reconstructed which leads to excellent background discrimination. The disadvantages are (1) normally the energy resolution is not good; (2) in order to release the beta decay electrons, the source has to be made into thin foils and hence large masses are difficult to integrate in an experiment.

The previous experiments include TGV I and II~\cite{Ste98, Ste00,   Ste06}, NEMO I~\cite{Das91}, II~\cite{Arn95} and III~\cite{Arn05,   Arn07}. In TGV Cd plates were put in between germanium spectrometers.  In NEMO the source foils (Ca, Se, Zr, Cd, Mo, Te and Nd) were fixed between two tracking volumes composed of many drift cells. The planned future experiments include SuperNEMO~\cite{Sne08}, MOON~\cite{Nak06} and DCBA~\cite{Ish05}. Based on the NEMO III experience, SuperNEMO aims at the sensitivity of $m_{\beta\beta} \sim 0.03$~eV. In MOON, enriched $^{100}$Mo foils will be interleaved with plastic scintillators which work as a calorimeter as well as an active shield.  The designed sensitivity of MOON is $m_{\beta\beta} \sim 0.03$~eV. In DCBA thin source plates ($^{150}$Nd,$^{100}$Mo $^{82}$Se) will be installed in tracking chambers located in a uniform magnetic field. DCBA is an R\&D project for MTD (Magnetic Tracking Detector), the design sensitivity of which is $m_{\beta\beta} \sim 0.02$-0.07~eV~\cite{Ish07}.

\subsection{Summary}
\label{sec:exp:comp}
Table~\ref{tab:gerda:comp} summarizes the proposed $0\nu\beta\beta$ decay experiments mentioned in the previous discussion. The schedule is not clear in all case. However, the table provides a quick reference concerning the experimental aspects of $0\nu\beta\beta$ decay research.

\begin{table}[htbp]
  \centering
  \caption{Characteristics of proposed $0\nu\beta\beta$ experiments.                         The corresponding references are: CARVEL~\cite{Zde05},     CANDLES~\cite{Hir08}, GERDA~\cite{Sch05,Cal06}, Majorana~\cite{Aal04},     Cobra~\cite{Kie03}, CAMEO~\cite{Bel01}, CUORE~\cite{Ard05},     XMASS~\cite{Nak02}, EXO~\cite{Dan00}, SNO+~\cite{Zub07},             MOON~\cite{Nak06}, SuperNEMO~\cite{Sne08}, DCBA/MTD~\cite{Ish07}}
  \label{tab:gerda:comp}
  \begin{minipage}{\linewidth}
    \begin{tabular}{lllc} \hline Experiment & Technique & Sensitivity       ($m_{\beta\beta}$/eV) & Schedule \\\hline
      CARVEL & CaWO$_{4}$  scintillator & 0.04-0.09 & - \\
      CANDLES & CaF$_{2}$ scintillator & - & - \\
      GERDA Phase I & Ge detector in LAr~\footnote{liquid argon} &       0.09-0.29 & 2009 \\
      GERDA Phase II & Ge detector in LAr & 0.05 & - \\
      Majorana & $^{76}$Ge detector & 0.03-0.04 & - \\
      Cobra & CdZnTe semiconductor & $< 1$ & - \\
      CAMEO & CdWO$_{4}$ scintillator & $< 0.02$ & - \\
      CUORE & TeO$_{2}$ bolometer & $< 0.03$ & - \\
      XMASS & liquid Xe TPC & 0.06-0.09 & - \\
      EXO & liquid Xe TPC with laser tagging & $< 0.01$ & - \\
      SNO+ & Nd-load scintillator & $< 0.05$ & 2010 \\ 
      MOON & Mo foil interleaved with scintillators & $\sim0.03$ & -\\
      SuperNEMO & drift chamber + calorimeter & $< 0.05$ & 2012 \\
      DCBA/MTD & source plates in drift chamber & 0.02-0.07 & 2016 \\
    \end{tabular}
  \end{minipage}
\end{table}

%%% Local Variables:
%%% mode:latex
%%% TeX-master: "thesis"
%%% End:
