\addcontentsline{toc}{chapter}{Introduction}
\chapter*{Introduction}
\label{cha:intro}
At the time the \emph{Standard Model} was established the neutrino was
believed to be massless. In experiments it always had the same
chirality and there was no evidence for a non-zero mass. However, the
picture changed dramatically when neutrino oscillations were observed
in solar and atmospheric neutrinos. They were explained by the weak
interaction eigenstates of neutrinos being admixtures of mass
eigenstates and the latter propagating with different velocities. The
introduction of neutrino mass terms into the Standard Model became
necessary.

There are various methods to introduce neutrino mass terms into the
Standard Model. The most straightforward approach is to follow the
same procedure as for the charged leptons; \textit{i.e.}, the leptons
obtain mass by coupling to the Higgs field. The problems of this
approach are that it does not explain why neutrinos couple to the
Higgs field so weakly compared to their charged partners, and that it
requires the introduction of right-handed neutrinos which have not yet
been experimentally observed. An elegant way to solve these problems
is to assume that neutrinos are Majorana particles; \textit{i.e.},
their own anti-particles. This way, the second problem does not arise,
and once the Majorana mass terms are introduced into the Lagrangian,
the so-called \emph{see-saw mechanism} can make the different coupling
strengths look natural.

The only experimental test currently possible to verify that neutrinos
are Majorana particles is the search for neutrinoless double beta
($0\nu\beta\beta$) decay. In this process, a neutrino emitted from one
neutron is absorbed as an anti-neutrino by another neutron. This can
only occur if neutrinos are of Majorana type. About ten naturally
occurring isotopes are observed to undergo double-beta decay. Among
them, $^{76}Ge$ is of special importance because germanium is a
semiconductor material used in highly efficient detectors with very
good energy resolution (it can serve as source and detector
simultaneously), and it is one of the purest materials produced in the
world limiting intrinsic background.

The GERDA (GERmanium Detector Array) experiment \cite{Abt04, Sch05},
searching for the $0\nu\beta\beta$ decay of $^{76}Ge$, is currently
under construction in Hall A of the INFN Gran Sasso National
Laboratory (LNGS), Italy. In order to achieve an extremely low
background level, it is considered to use 18-fold segmented germanium
detectors operated directly in cryogenic liquid, serving as cooling
and shielding material, in the second phase of GERDA. The main goals
of the work presented in this thesis are to examine systematically the
operation and performance of segmented detectors in cryogenic liquid
and to investigate their power of background discrimination by
analyzing the spatial distribution over which energy is deposited. The
time structure of the detector response is studied and simulated to
lay the foundation for its use in background suppression.

Two test facilities were used to take data with two 18-fold segmented
GERDA Phase~II prototype detectors:\\
\emph{Gerdalinchen~II}, a specially designed cryostat containing
liquid nitrogen or argon, inside which up to three segmented detectors
can be operated simultaneously. It was used
\begin{itemize}
\item to operate and study for the first time segmented detectors
submerged directly in cryogenic liquid.
\item to develop detailed operating procedures.
\item to carefully study the data of a prototype detector.
\end{itemize}
\emph{A vacuum cryostat}, especially equipped to operate one segmented
detector. It allowed
\begin{itemize}
\item the analysis of background induced by external photons in the
MeV-energy range. These photons typically undergo multiple Compton
scattering and deposit their energies over a range of several
centimeters. This distinguishes them from the electrons from
$0\nu\beta\beta$ decay which deposit energy on a millimeter scale.
\item the analysis of background induced by neutron interactions with
germanium isotopes and surrounding materials. Most of the neutron
induced events deposit their energies in several segments of the
detector. In particular, the inelastic scattering of neutrons on
germanium isotopes can be identified through the separation of the
energies deposited by the prompt photon and the nuclear recoil.
\item the characterization of the detector, including the segment
boundaries, crystal axes, impurities, \textit{etc.}. The data was used
to study the time structure of the detector response, \textit{i.e.},
the pulse shape and to verify the corresponding simulations.
\end{itemize}

The test facilities were modeled using a Geant4 based simulation
framework, MaGe, which is jointly developed by the GERDA and Majorana
collaborations. The simulation of low energy electrons, photons and
neutrons interacting with germanium detectors and surrounding
materials were verified in detail. A fully functional pulse shape
simulation package was also developed within the MaGe framework. The
whole signal formation process in segmented germanium detectors and
the read-out system was simulated and verified by being compared with
data.

The thesis is structured as follows:
\begin{description}
\item[Chapter~\ref{cha:theory}] describes the theoretical background
for $0\nu\beta\beta$ decay as well as other approaches to test whether
neutrinos are of Majorana or Dirac type.
\item[Chapter~\ref{cha:exps}] summarizes different technical
approaches for searching for $0\nu\beta\beta$ decays of different
isotopes, compares the experiments with each other and estimates the
potential of future $0\nu\beta\beta$ decay experiments.
\item[Chapter~\ref{cha:gerda}] introduces the basic ideas of the GERDA
experiment, summarizes the latest progress, and estimates the
potential of GERDA.
\item[Chapter~\ref{cha:detector}] describes the basic concepts of
semiconductor detectors and the important properties of germanium
crystals and detectors related to the later analysis.
\item[Chapter~\ref{cha:teststand}] introduces the two test stands that
provided the data for all studies and describes the slow control and
data acquisition system on which the test stands rely.
\item[Chapter~\ref{cha:GII}] characterizes the short and long term
performance of segmented germanium detectors submerged in cryogenic
liquid.
\item[Chapter~\ref{cha:np}] describes a special class of events with
negative baseline shifts.
\item[Chapter~\ref{cha:photon}] demonstrates the power of segmented
detectors to identify photon induced background and verifies the Monte
Carlo simulation.
\item[Chapter~\ref{cha:neutron}] demonstrates the power of segmented
detectors to identify neutron interactions with germanium isotopes and
surrounding materials and validates the simulation in this aspect as
well.
\item[Chapter~\ref{cha:pss}] describes the physics models of the
charge carrier drift inside germanium detectors to simulate the pulse
shape and introduces methods to add electronic effects.
\item[Chapter~\ref{cha:psa}] verifies the pulse shape simulation by
comparing it to the data taken with the GERDA prototype detector and
introduces new methods to determine the crystal orientation and
impurity distributions.
\end{description}
The results are summarized within the context of the GERDA Phase~II
experiment and an outlook to further studies is given.


%%% Local Variables:
%%% mode:latex
%%% TeX-master: "thesis"
%%% End:
